---
title: "Multiple Comparisons - Benjamini-Hochberg Procedure and Bonferroni Correction"
author: Naomi Johnson
output: html_notebook
---

Exercise: 
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 10  # Mean of sample 1
mean2 <- 10  # Mean of sample 2
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Print the first 10 p-values
cat("First 10 p-values:\n")
cat(p_values[1:10], sep = ", ")

# Save p-values to a CSV file
results_df <- data.frame(P_Value = p_values)
write.csv(results_df, "t_test_p_values.csv", row.names = FALSE)


```

Applying the Bonferroni Correction
```{r}
# Load the saved p-values
p_values <- read.csv("t_test_p_values.csv")$P_Value

# Number of tests conducted
num_tests <- length(p_values)

# Apply Bonferroni correction
bonferroni_corrected_p_values <- p_values * num_tests

# Print the first 10 Bonferroni-corrected p-values
cat("First 10 Bonferroni-corrected p-values:\n")
cat(bonferroni_corrected_p_values[1:10], sep = ", ")

```
Applying the Benjamini-Hochberg Procedure
```{r}
# Load the saved p-values
p_values <- read.csv("t_test_p_values.csv")$P_Value

# Apply the Benjamini-Hochberg procedure
bh_adjusted_p_values <- p.adjust(p_values, method = "BH")

# Print the first 10 BH-adjusted p-values
cat("First 10 BH-adjusted p-values:\n")
cat(bh_adjusted_p_values[1:10], sep = ", ")

```
Re-run the exercise with sample 1 mean = 1 and sample 2 mean = 2
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 2  # Mean of sample 2
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Print the first 10 p-values
cat("First 10 p-values:\n")
cat(p_values[1:10], sep = ", ")

# Save p-values to a CSV file
results_df <- data.frame(P_Value = p_values)
write.csv(results_df, "t_test_p_values_mean_diff.csv", row.names = FALSE)

```
Re-run with the difference in means even greater (in this case, the difference is 5)
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 6  # Mean of sample 2 (larger difference)
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Print the first 10 p-values
cat("First 10 p-values:\n")
cat(p_values[1:10], sep = ", ")

```
Applying the Bonferroni Correction (for sample 1 with mean = 1, sample 2 with mean = 2)
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 2  # Mean of sample 2
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Apply Bonferroni correction
bonferroni_corrected_p_values <- p_values * num_simulations

# Print the first 10 Bonferroni-corrected p-values
cat("First 10 Bonferroni-corrected p-values:\n")
cat(bonferroni_corrected_p_values[1:10], sep = ", ")

```
Applying the Benjamini-Hochberg Procedure
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 2  # Mean of sample 2
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Apply Benjamini-Hochberg correction
bh_adjusted_p_values <- p.adjust(p_values, method = "BH")

# Print the first 10 BH-adjusted p-values
cat("First 10 BH-adjusted p-values:\n")
cat(bh_adjusted_p_values[1:10], sep = ", ")

```
Applying the Bonferroni Correction (for sample 1 with mean = 1, sample 2 with mean = 6)
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 6  # Mean of sample 2 (larger difference)
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Apply Bonferroni correction
bonferroni_corrected_p_values <- p_values * num_simulations

# Print the first 10 Bonferroni-corrected p-values
cat("First 10 Bonferroni-corrected p-values:\n")
cat(bonferroni_corrected_p_values[1:10], sep = ", ")

```
Applying the Benjamini-Hochberg
```{r}
# Set the seed for reproducibility
set.seed(123)

# Define parameters
sample_size <- 30  # Size of each sample
num_simulations <- 1000  # Number of simulations
mean1 <- 1  # Mean of sample 1
mean2 <- 6  # Mean of sample 2 (larger difference)
sd1 <- 2     # Standard deviation of sample 1
sd2 <- 2     # Standard deviation of sample 2

# Initialize a vector to store p-values
p_values <- numeric(num_simulations)

# Perform 1000 t-tests and store p-values
for (i in 1:num_simulations) {
  # Generate two random samples
  sample1 <- rnorm(sample_size, mean = mean1, sd = sd1)
  sample2 <- rnorm(sample_size, mean = mean2, sd = sd2)
  
  # Perform a t-test and store the p-value
  t_result <- t.test(sample1, sample2)
  p_values[i] <- t_result$p.value
}

# Apply Benjamini-Hochberg correction
bh_adjusted_p_values <- p.adjust(p_values, method = "BH")

# Print the first 10 BH-adjusted p-values
cat("First 10 BH-adjusted p-values:\n")
cat(bh_adjusted_p_values[1:10], sep = ", ")

```
Observations: The p-values decrease with increasing differences in the mean for the two samples. 
